---
title: "O2 Target Meta-Analysis"
output: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Oxygen targets in ICU patients

This workbook synthesizes the available evidence on oxygen targets in the ICU. 

The aim of this project is to take the 'outside view' of the data to predict the results and interpretation of the ongiong Mega-Rox trial.

Note: my presupposition is that there is not relative effect modification 

```{r}
#| eval: false
#| label: Pull in Data Extraction Spreadsheet
library(readxl)
library(tidyverse)
# Mac Version
setwd("/Users/blocke/Box Sync/Residency Personal Files/Stats/O2 Target Meta-Analysis") 

# PC Version
#setwd('C:\\Users\\reblo\\Box\\Residency Personal Files\\Stats\\O2 Target Meta-Analysis\\') #[ ] figure out how to make this a global

data_sheet <- read_excel("data/O2 icu target MA.xls") |> 
  select(-other_comment) |>  # drop other comment column
  arrange(year) # arrange by year (needed for TSA) 

head(data_sheet)
authors <- select(data_sheet, author)
```


## Data Extraction tables

[ ] todo: 
please double check me if any of my extractions are erroneous.

## Meta-analysis

### Random effects meta-analysis using the meta package

```{r}
library(meta)

#metabin takes events, total (rather than events, nonevents)
m_ex1 <- meta::metabin(low_o2_died, num_low_o2, high_o2_died, num_high_o2, data = data_sheet, studlab = paste(name, author, year), sm = "OR")
meta::forest(m_ex1, comb.random = FALSE, lab.c = "High Oxygen", lab.e = "Low Oxygen", label.left = "Favors Low O2", label.right = "Favors High O2")

```

### Random effects meta-analysis using the metafor package

Using Mantel Haenszel Estimator

```{r}
library(metafor)
#this one takes the events and non-events
m4_ex1 <- rma.mh(low_o2_died, low_o2_alive, high_o2_died, high_o2_alive, data = data_sheet, slab = paste(author, year, sep = ", "))
forest(m4_ex1, transf = exp, showweights = TRUE, refline = 1, header = "Author and Year", shade = "zebra", ilab=cbind(low_o2_died, low_o2_alive, high_o2_died, high_o2_alive))

```

### Funnel plot

Nothing in particular suspicious about this
```{r}
funnel(m_ex1) # this one will take either metafor or meta object

metabias(m_ex1, k.min=5) # note: this one takes meta, not metafor object

```

## Trial Sequential Analysis

Note: this analysis remains finicky for reasons I don't understand yet. 

0However, running for an expected 10% RRR, we can say that futility has already been met (ie, further data is unlikely to the usual 5% type 1 error rate to change the assessment that there is not a 10% RRR)

However, much more likely - if there is an effect it is much, much smaller (e.g. 1-3%), but that this would still be HUGELY important, as realistically there is no known downside to a different oxygen target within the range being tested, and thus ANY real mortality benefit would be important. Hence, the MEGA-ROX trial is important and warranted. 

```{r}
library(RTSA)
# library(ldbounds)


# Have to rename columns such that there is: 
# ‘eI‘(eventsininterventiongroup), ‘eC‘(eventsincontrolgroup),‘nC‘(participantsinterventiongroup)or‘nI‘ (participantscontrolgroup)fordiscretedata

rtsa_df <- data_sheet |> 
  rename("eI" = high_o2_died, 
         "eC" = low_o2_died, 
         "nI" = num_high_o2,
         "nC" = num_low_o2, 
         "study" = name)

map(rtsa_df, class)
sapply(rtsa_df, function(x) which(is.na(x)))


an_rtsa <- RTSA(type="analysis", data =  rtsa_df , outcome = "RR", mc = 0.9, side = 2,  alpha = 0.05, beta = 0.1, es_alpha = "esOF", es_beta = "esOF", futility = "non-binding", random_adj = "D2", D2 = 0.25 )

#an_rtsa <- RTSA(type = "analysis", data = rtsa_df, outcome = "OR", side = 2, es_alpha = "esOF", es_beta = "esOF", fixed = FALSE, re_method = "DL_HKSJ", alpha = 0.05, beta = 0.1, random_adj = "D2", D2 = 0.25, mc = 0.02)
plot(an_rtsa)


```


> setwd("C:\\Users\\reblo\\Box\\Residency Personal Files\\MSCI\\MDCRC 6200 SRMA\\O2 TargetMeta-anlysis")
> t <- c(.0543122962117195, .0672617554664612, .1218254938721657, .1299974918365479, .1450842320919037, .2664068341255188, .2916771471500397
> , .6547648906707764, .7050540447235107, .9165199995040894, 1)
> obf.bd <- ldBounds( t, iuse = 1, alpha = .025) 
> obf = data.frame( t_bd = obf.bd$time, UB_bd = obf.bd$upper.bounds)
> write.dta(obf,"bounds.dta")


## Network Meta-analysis


## Expected Results of the MEGA-ROX trial. 

