---
title: "NRH Vent Weaning"
author: "Brian Locke"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Data Cleaning

Import needed utilities and the data to a tbl named df 

```{r import data}
#| label: Pull in Data Extraction Spreadsheet
#| echo: false
library(readxl)
library(ggplot2)
library(tidyverse) # todo: import each of these individually to avoid the error message
setwd("/Users/reblocke/Box Sync/Residency Personal Files/Stats/NRH Vent Proj") #[ ] TODO: figure out how to make this a global
df_orig <- read_excel("data/2023-8-22 NRH SCI FINAL.xlsx")
```

Drop un-needed columns for current analysis: 
```{r drop columns}
df <- df_orig |>
  select(-PAT_MRN_ID, -`Date of Injury`, -Column1, -Comments)
#Getting rid of date for now, though might be of use later - would need to figure out how to convert the dates. 
```

Rename variables to usual R syntax
```{r rename vars}
df <- df |>
  rename(age = Age, 
         female = Sex,
         prior_inj_level = `Injury Level reported prior to rehab`,
         rib_fx = `Rib fractures (Y or N)`,
         ptx = `Pneumothorax (Y or N)`,
         chest_tube = `Chest tube (Y or N)`,
         days_inj_intub = `Days from injury to intubation`,
         reason_intub = `Reason for Intubation`,
         days_intub_trach = `Days from intubation to Trach`, 
         wean_before_trach = `Attempt to wean prior to Trach. (Y or N)`,
         wean_before_transfer = `Attempt to wean off vent prior to transfer to rehab (Y or N)`,
         pna_before_transfer = `Did patient develop pneumonia prior to rehab admission (Y or N)`,
         admit_from_osh = `OSH?`,
         admit_from = `Admit From`,
         days_inj_rehab = `Days from injury to admission to rehab`,
         asia_class_at_rehab = `ASIA Classification at Rehab`,
         noc_vent_only = `Wean off vent during the day and continue to need nightime ventilation? (Y or N)`,
         days_adm_to_noc_vent_only = `If yes to previous question, How many days from admission to rehab until they weaned?`,
         full_wean = `Wean off vent for full 24 hours? (Y or N)`,
         days_adm_to_full_wean = `If yes to previous question, How many days from admission to rehab until they weaned? 2`,
         decannulated = `Did patient decanulate (Y or N)`,
         days_adm_to_decan = `Days from admission to rehab to decanulation`,
         transfer_icu_rehab = `Did they transfer to ICU from rehab. (Y or N)`,
         transfer_icu_rehab_resp = `If they transferred to ICU, was cause respirtory?  (Y or N)`,
         pna_at_rehab = `Did they Develop Pnumonia at rehab? (Y or N)`,
         days_adm_disch_rehab = `Days to discharge from rehab`,
         disch_loc = `Discharge Location`,
         rec_complete = `Complete record`)
```

See what the current class of each columnn is
```{r}
map(df, class)
```

Define programs to clean variable types: 
Clean all the numerical values
Clean all the yes or no binary values 

```{r helpers}
clean_numerical <- function(value) { 
  cleaned <- map_int(value)
  return(cleaned)
}

# takes the yes/no and makes them boolean
# works by chopping off everything after the first letter
clean_yesno <- function(value) { 
  # These lines turn 'NA' that are in the doc into Un, which starts with U, rather than N = NO"
  #value <- substring(value, 1, 2) |> 
  #  recode(`NA` = "Un", 
  #         `na`="Un")
  cleaned <- substring(value, 1, 1) |> 
    recode(Y = TRUE,
           y = TRUE,
           N = FALSE,
           n = FALSE,
           u = NA,
           U = NA,
           )
  return(cleaned)
  }
```


Recode all of the columns to their maximally specific data-types

```{r recoding}
# Recode M's and F's to boolean logic
df$female <- df$female |> 
  recode(male = FALSE, M = FALSE, `F` = TRUE)

# TODO: split level of injury logic; for now just left as strings. 

# Change the following columns from (yes/no/unknown) to (TRUE/FALSE/NA)
df$rib_fx <- clean_yesno(df$rib_fx)
df$ptx <- clean_yesno(df$ptx)
df$chest_tube <- clean_yesno(df$chest_tube)
df$wean_before_trach <- clean_yesno(df$wean_before_trach)
df$wean_before_transfer <- clean_yesno(df$wean_before_transfer)
df$pna_before_transfer <- clean_yesno(df$pna_before_transfer)
df$admit_from_osh <- clean_yesno(df$admit_from_osh)
df$noc_vent_only <- clean_yesno(df$noc_vent_only)
df$full_wean <- clean_yesno(df$full_wean)
df$decannulated <- clean_yesno(df$decannulated)
df$transfer_icu_rehab <- clean_yesno(df$transfer_icu_rehab)
df$transfer_icu_rehab_resp <- clean_yesno(df$transfer_icu_rehab_resp)
df$pna_at_rehab <- clean_yesno(df$pna_at_rehab)
df$rec_complete <- clean_yesno(df$rec_complete)

#Change the following into numerical variables
df$days_adm_to_noc_vent_only <- parse_number(df$days_adm_to_noc_vent_only) # note, for this one - there are a few that are 'prior', that are treated as NA... might also consider 0. 
df$days_adm_to_full_wean <- parse_number(df$days_adm_to_full_wean) #TODO: capture the "1"

```



TODO: do the categorical variables nee to be explicitly recoded into a categorical variables, as in STATA?  Below is the relevant STATA code


encode injurylevelreportedpriortorehab, gen(init_injury_level)
drop injurylevelreportedpriortorehab
label variable init_injury_level "Initial Injury Level"

replace asiaclassificationatrehab = strtrim(strupper(asiaclassificationatrehab))
split asiaclassificationatrehab, gen(level_class)
replace level_class2 = "AIS" if level_class2 == "AIA"
encode level_class1, generate(level)
label variable level "Level of Injury"
drop level_class1
generate asia_class_temp = level_class2 + " " + level_class3
encode asia_class_temp, gen(asia_class)
drop level_class2 level_class3 asia_class_temp asiaclassificationatrehab
label variable asia_class "ASIA classification"

generate high_v_low = ""
//replace high_v_low = "High" if level == "C1"
replace high_v_low = "High" if level == 1
replace high_v_low = "High" if level == 2
replace high_v_low = "High" if level == 3
replace high_v_low = "Low" if level == 4
replace high_v_low = "Low" if level == 5
replace high_v_low = "Low" if level == 6
encode high_v_low, gen(high_vs_low)
drop high_v_low
label variable high_vs_low "High (C4 or above) or Low (C5 or below) Cervical Injury?"

generate init_high_v_low = ""
replace init_high_v_low = "High" if init_injury_level == 1
replace init_high_v_low = "High" if init_injury_level == 2
replace init_high_v_low = "High" if init_injury_level == 3
replace init_high_v_low = "High" if init_injury_level == 4
replace init_high_v_low = "Low" if init_injury_level == 5
replace init_high_v_low = "Low" if init_injury_level == 6
replace init_high_v_low = "Low" if init_injury_level == 7
encode init_high_v_low, gen(init_high_vs_low)
drop init_high_v_low
label variable init_high_vs_low "Initially Assessed as High (C4 or above) or Low (C5 or below) Cervical Injury?"

gen weaning_outcome_str = ""
replace weaning_outcome_str = "Fully Vent Dependent" if wean_during_day == 1 
replace weaning_outcome_str = "Weaned During Day" if wean_during_day == 2
replace weaning_outcome_str = "Fully Weaned Off" if wean_24hr == 2
replace weaning_outcome_str = "Decannulated" if decannulate == 2
label define weaning_label 1 "Fully Vent Dependent" 2 "Weaned During Day" 3 "Fully Weaned Off" 4 "Decannulated"
encode weaning_outcome_str, label(weaning_label) gen(weaning_outcome)
drop weaning_outcome_str
label variable weaning_outcome "Ordinal outcome of weaning attempts"

label define discharge_label 1 "HOME" 2 "HOME W/ HOME HEALTH" 3 "SNF" 4 "LTAC"
replace dischargelocation = strtrim(strupper(dischargelocation))
replace dischargelocation = "HOME W/ HOME HEALTH" if dischargelocation == "HOME HEALTH"
replace dischargelocation = "HOME W/ HOME HEALTH" if dischargelocation == "HH"
replace dischargelocation = "LTAC" if dischargelocation == "NEURORESTORITIVE THERAPY"
encode dischargelocation, label(discharge_label) gen(discharge_to)
label variable discharge_to "Discharge Location?"
drop dischargelocation

```{r}
#TODO

```



Confirm that each column has been forced to the correct columns and compare to the original data
```{r data cleaning checks}

map(df, class)
map(df, n_distinct) # summarize number of distinct entries for each

# TODO: ensure that these are all working? compare to original perhaps

```

## Compute summary statistics for the data 

STATA summary statistics
tab male
summarize age, detail

//ASIA score? AIS? for cross tabulation.
//Injury level vs level?
summarize daysfrominjurytointubation, detail // immediate vs not-immediate as useful dichotomy to evaluate.
summarize daysfromintubationtotrach, detail
summarize daysfrominjurytoadmissiontorehab, detail
summarize days_to_daytime_wean, detail
summarize days_to_24hr_wean, detail
summarize daystodischargefromrehab, detail
tab level asia_class
tab level init_injury_level

```{r}
#TODO
```

## Create tables of the data 

Example STATA code to replicate
table1_mc, by(level) ///
vars( ///
male cat %4.0f \ ///
age conts %4.0f \ ///
rib_fx cat %4.0f \ ///
ptx cat %4.0f \ ///
chest_tube cat %4.0f \ ///
init_injury_level cat %4.0f \ ///
asia_class cat %4.0f \ ///
daysfrominjurytointubation conts %4.0f \ ///
wean_pre_trach cat %4.0f \ ///
daysfromintubationtotrach conts %4.0f \ ///
wean_pre_trans cat %4.0f \ ///
daysfrominjurytoadmissiontorehab conts %4.0f \ ///
pneumonia_prior cat %4.0f \ ///
wean_during_day cat %4.0f \ ///
days_to_daytime_wean conts %4.0f \ ///
wean_24hr cat %4.0f \ ///
days_to_24hr_wean conts %4.0f \ ///
decannulate cat %4.0f \ ///
weaning_outcome cat %4.0f \ ///
pna_at_rehab cat %4.0f \ ///
rehab_to_icu cat %4.0f \ ///
resp_icu_transfer cat %4.0f \ ///
discharge_to cat %4.0f \ ///
daystodischargefromrehab conts %4.0f \ ///
) ///
percent_n percsign("%") iqrmiddle(",") sdleft(" (±") sdright(")") onecol total(before) ///
saving("Results and Figures/$S_DATE/All by Level.xlsx", replace)


table1_mc, by(asia_class) ///
vars( ///
male cat %4.0f \ ///
age conts %4.0f \ ///
rib_fx cat %4.0f \ ///
ptx cat %4.0f \ ///
chest_tube cat %4.0f \ ///
init_injury_level cat %4.0f \ ///
level cat %4.0f \ ///
asia_class cat %4.0f \ ///
daysfrominjurytointubation conts %4.0f \ ///
wean_pre_trach cat %4.0f \ ///
daysfromintubationtotrach conts %4.0f \ ///
wean_pre_trans cat %4.0f \ ///
daysfrominjurytoadmissiontorehab conts %4.0f \ ///
pneumonia_prior cat %4.0f \ ///
wean_during_day cat %4.0f \ ///
days_to_daytime_wean conts %4.0f \ ///
wean_24hr cat %4.0f \ ///
days_to_24hr_wean conts %4.0f \ ///
decannulate cat %4.0f \ ///
weaning_outcome cat %4.0f \ ///
pna_at_rehab cat %4.0f \ ///
rehab_to_icu cat %4.0f \ ///
resp_icu_transfer cat %4.0f \ ///
discharge_to cat %4.0f \ ///
daystodischargefromrehab conts %4.0f \ ///
) ///
percent_n percsign("%") iqrmiddle(",") sdleft(" (±") sdright(")") onecol total(before) ///
saving("Results and Figures/$S_DATE/All by Class.xlsx", replace)


tab level init_injury_level

table1_mc, by(init_high_vs_low) ///
vars( ///
male cat %4.0f \ ///
age conts %4.0f \ ///
rib_fx cat %4.0f \ ///
ptx cat %4.0f \ ///
chest_tube cat %4.0f \ ///
init_injury_level cat %4.0f \ ///
level cat %4.0f \ ///
daysfrominjurytointubation conts %4.0f \ ///
wean_pre_trach cat %4.0f \ ///
daysfromintubationtotrach conts %4.0f \ ///
wean_pre_trans cat %4.0f \ ///
daysfrominjurytoadmissiontorehab conts %4.0f \ ///
pneumonia_prior cat %4.0f \ ///
wean_during_day cat %4.0f \ ///
days_to_daytime_wean conts %4.0f \ ///
wean_24hr cat %4.0f \ ///
days_to_24hr_wean conts %4.0f \ ///
decannulate cat %4.0f \ ///
weaning_outcome cat %4.0f \ ///
pna_at_rehab cat %4.0f \ ///
rehab_to_icu cat %4.0f \ ///
resp_icu_transfer cat %4.0f \ ///
discharge_to cat %4.0f \ ///
daystodischargefromrehab conts %4.0f \ ///
) ///
percent_n percsign("%") iqrmiddle(",") sdleft(" (±") sdright(")") onecol total(before) ///
saving("Results and Figures/$S_DATE/All by Init High vs Low.xlsx", replace)

```{r}
#TODO
```

## Regression analyses: 

Example STATA code to replace
//TODO: integrate info on where they came from
// levels vs ASIA can differ because ASIA is based on functionality and can evolve.
// how often do the labels change? do they go up or down? does grade matter? 
// ASIA -->  true assessment
// some of the difference is less experienced examiners, some of it is spinal shock.
// impact of this: would argue we should wait for prognostic assessments and communications to patients.
 
//Exposures: where they came from (UU vs not; SICU vs NCCU vs MICU) ; aggressiveness of weaning
//Confounders: age, sex, severity, level, time to admission to rehab

//STATA regressino code to replicate
ologit discharge c.age c.daysfrominjurytoadmissiontorehab ib3.level, or
ologit discharge c.age c.daysfrominjurytoadmissiontorehab ib1.asia_class, or

ologit discharge i.wean_pre_trach c.age ib3.level, or
ologit discharge i.wean_pre_trach c.age ib1.asia_class, or

ologit discharge i.wean_pre_trans c.age ib3.level, or
ologit discharge i.wean_pre_trans c.age ib1.asia_class, or

ologit weaning_outcome c.age c.daysfrominjurytoadmissiontorehab ib3.level, or
ologit weaning_outcome  c.age c.daysfrominjurytoadmissiontorehab ib1.asia_class, or

ologit weaning_outcome i.wean_pre_trach c.age ib3.level, or
ologit weaning_outcome i.wean_pre_trach c.age ib1.asia_class, or

ologit weaning_outcome i.wean_pre_trach c.daysfrominjurytoadmissiontorehab c.age i.high_vs_low, or

## Visualizations

General Visualizations

```{r}
#| label: figure-1
#| fig-cap: "Timing of Trach and Intubation"

#note; fig-cap will only appear if you give a label option. 

df |> 
  ggplot(mapping = aes(x = days_inj_intub, 
                       y = days_intub_trach,
                       colour = pna_before_transfer)) + 
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Days between injury to intubation (log)",
    y = "Days between intubation and trach (log)", 
    color = "Pna before transfer?", 
    title = "Timing of intubation and trach by pneumonia"
  )

```


```{r age-by-los}

df |> 
  ggplot(mapping = aes(x = age, 
                       y = days_adm_disch_rehab,
                       colour = asia_class_at_rehab)) + 
  scale_y_log10() +
  geom_point(alpha = 0.5) +
  labs(
    x = "Age (years)",
    y = "Length of Stay (log)", 
    color = "ASIA scale?", 
    title = "Age vs LOS at Rehaab") +
  facet_wrap(~female)

```

TODO:  Sankey Diagrams? 


//plan: was there internal referral? Is this a way to make pre-admission weaning comparable?
//do people who wean faster lower the total length of stay
// how frequent was pneumonia, controlling for pneumonia that occured before. 

// Outcomes: 
// Ordinal: decannulate, weaned 24h, wean during day, no wean     [1 limitation of trach not vented: could be placement to facility or TBI]
// Ordinal: discharge location ****
// Cox-regression for length of stay.

//Old data: asia class -> likelihood to wean 

//how could you generate useful preliminary data to support an early aggressive weaning protocol. -- implicit comparison to historical rates of weaning?
//RT flow sheets for data on what is happening. 
