---
title: "sci vent chart study"
format: html
editor: visual
---

## Demographics

```{r, echo=FALSE, message=FALSE}
library(dplyr)
sci_cleanned <- readRDS("sci_cleanned.rds")
colnames(sci_cleanned)[colnames(sci_cleanned) == 'Wean.off.vent.for.full.24.hours.'] <- '24_hours_off_vent'
colnames(sci_cleanned)[colnames(sci_cleanned) == 'If.yes.to.previous.question..How.many.days.from.admission.to.rehab.until.they.weaned..'] <- 'days_to_wean_during_day'
colnames(sci_cleanned)[colnames(sci_cleanned) == 'Days.to.wean.24.hours'] <- 'days_to_wean_24hours'
row_to_change <- 19
sci_cleanned$'24_hours_off_vent'[row_to_change] <- FALSE
```

```{r, echo=FALSE, results="asis", message=FALSE}
Level_table <- sci_cleanned %>% 
  summarise(C2 = sum(level == "C2"),
            C3 = sum(level == "C3"),
            C4 = sum(level == "C4"),
            C5 = sum(level == "C5"),
            C6 = sum(level == "C6"),
            C7 = sum(level == "C7"))
demographics_table <- sci_cleanned %>% 
  summarise(Male = sum(Sex == "M"),
            Female = sum(Sex == "F"),
            'Mean Age' = round(mean(Age),2),)
asia_table <- sci_cleanned %>% 
  select(level, asia) %>% 
  group_by(level, asia) %>% 
  summarise(
    n=n()
  )

print(knitr::kable(demographics_table))
print(knitr::kable(Level_table))
print(knitr::kable(asia_table))
```

## Days from intury to intubation and trach grouped by injury level

```{r, echo=FALSE, results="asis", message=FALSE}
library(dplyr)
results_by_level <- sci_cleanned %>% 
  select(level, injury.to.intubation, intubation.to.Trach) %>% 
  group_by(level) %>% 
  summarise(
    n=n(),
    days_from_inj_to_int= round(mean(injury.to.intubation), 2), 
    days_from_int_to_trach = round(mean(intubation.to.Trach), 2)
  )


print(knitr::kable(results_by_level))
```

## Count and percentage of patients who weaned and decanulated grouped by level and asia

```{r, echo=FALSE, results="asis", message=FALSE}
library(dplyr)
selected_columns <- sci_cleanned[c('PAT_MRN_ID', 'level', 'asia', 'off.vent.during.the.day', '24_hours_off_vent', 'decanulate', 'days_to_wean_during_day', 'days_to_wean_24hours', 'Days.to.decanulation', 'Days.to.discharge')]
asia_dataset <- data.frame(selected_columns)
asia_dataset$off.vent.during.the.day <- as.logical(asia_dataset$off.vent.during.the.day)
asia_dataset$ X24_hours_off_vent <- as.logical(asia_dataset$ X24_hours_off_vent)
asia_dataset$decanulate <- as.logical(asia_dataset$decanulate)

  
result_by_asia <- asia_dataset %>%
  group_by(level, asia) %>%
  summarize(
    n=n(),
    count_off_vent_day = sum(off.vent.during.the.day == TRUE),
    percentage_off_vent_day = round(mean(off.vent.during.the.day == TRUE) * 100, 2),
    count_off_vent_24_hours = sum(`X24_hours_off_vent` == TRUE),
    percentage_off_vent_24_hours = round(mean(`X24_hours_off_vent` == TRUE) * 100, 2),
    count_decanulate = sum(decanulate == TRUE),
    percentage_decanulate = round(mean(decanulate == TRUE) * 100, 2))

  
print(knitr::kable(result_by_asia))

```

## Average days from admission to rehab to wean off daytime vent

```{r, echo=FALSE, results="asis", message=FALSE}

wean_day <- sci_cleanned %>% 
  select(level, off.vent.during.the.day, days_to_wean_during_day) %>% 
  filter(off.vent.during.the.day == TRUE) %>%
  filter(!is.na(days_to_wean_during_day)) %>%
  group_by(level) %>% 
  summarise(
    n=n(),
    average_days_to_wean_day = round(mean(days_to_wean_during_day), 2))
  
print(knitr::kable(wean_day))   
```

## Average days from admission to rehab to wean 24 hours grouped by level

```{r, echo=FALSE, results="asis", message=FALSE}
wean_24hours <- sci_cleanned %>%
  select(level, `24_hours_off_vent`, days_to_wean_24hours) %>% 
  filter(`24_hours_off_vent` == TRUE) %>% 
  group_by(level) %>%
  summarise(
    n=n(), 
    average_days_to_wean_24hours = round(mean(days_to_wean_24hours), 2))
print(knitr::kable(wean_24hours))  
```

## Average days from admission to rehab to decanulation grouped by level

```{r, echo=FALSE, results="asis", message=FALSE}
decanulate_mean <- sci_cleanned %>% 
  select(level, decanulate, Days.to.decanulation) %>%
  filter(decanulate == TRUE) %>% 
  group_by(level) %>% 
  summarise(
    n=n(), 
    average_days_to_decanualtion = round(mean(Days.to.decanulation),2)
  )
print(knitr::kable(decanulate_mean))  
```

## Average days to discharge from rehab grouped by level

```{r, echo=FALSE, results="asis", message=FALSE}
average_days_to_discharge <- sci_cleanned %>%
  select(level, Days.to.discharge) %>%
  group_by(level) %>%
  summarize(
    n=n(), 
    average_days_discharge_reahb = round(mean(Days.to.discharge),2)
  )

print(knitr::kable(average_days_to_discharge)) 
```

## OSH time in days to intubation and intubation to trach.

```{r, echo=FALSE, results="asis"}

## OSH

osh_true <- sci_cleanned %>%
  filter(OSH. == TRUE) %>%
  group_by(level) %>%
  summarise(
    n=n(), 
    injury_to_int = round(mean(injury.to.intubation), 2), 
    int_to_trach = round(mean(intubation.to.Trach), 2), 
  )

print(knitr::kable(osh_true))

```

## U of U time in days to intubation and intubation to trach.

```{r, echo=FALSE, results="asis"}
osh_false <- sci_cleanned %>% 
  filter(OSH. == FALSE) %>% 
  group_by(level) %>% 
  summarise(
    n=n(),
    injury_to_int = round(mean(injury.to.intubation), 2), 
    int_to_trach = round(mean(intubation.to.Trach), 2),
  )

print(knitr::kable(osh_false))
```

## OSH vs UofU in days

```{r, echo=FALSE, results="asis"}
osh_true_total <- sci_cleanned %>% 
  filter(OSH.== TRUE) %>%
  summarise(
    n=n(),
        injury_to_int = round(mean(injury.to.intubation), 2), 
    int_to_trach = round(mean(intubation.to.Trach), 2),
  )

osh_false_total <- sci_cleanned %>% 
  filter(OSH.== FALSE) %>% 
  summarise(
    n=n(),
        injury_to_int = round(mean(injury.to.intubation), 2), 
    int_to_trach = round(mean(intubation.to.Trach), 2),
  )

osh_false_total$admission<- "U of U"
osh_true_total$admission <- "OSH"
osh_vs_uofu <- bind_rows(osh_false_total, osh_true_total) %>%
  select(admission, n, injury_to_int, int_to_trach)


print(knitr::kable(osh_vs_uofu))
```

```{r, echo=FALSE}
confounding_factors_true <- sci_cleanned %>% 
  filter(Rib.fractures == TRUE | Pneumothorax == TRUE | Chest.tube == TRUE) 

confounding_factors_false <- sci_cleanned %>% 
  filter(Rib.fractures == FALSE & Pneumothorax == FALSE & Chest.tube == FALSE) 
 
  
```
